Binocular combination in the autonomic nervous system

### 2024-11-28 manuscript.Rmd
Added back line numbers
Modifeid manuscript

---
title: 'Binocular combination in the autonomic nervous system'
author: |
 | Federico G. Segala$^1$, Aurelio Bruno$^{1,2}$, Joel T. Martin$^{1,3}$,
 | Anisa Y. Morsi$^1$, Alex R. Wade$^{1,4}$ & Daniel H. Baker$^{1,4}$
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: true
  html_document: default
bibliography: references.bib
link-citations: yes
linkcolor: black
csl: the-journal-of-neuroscience.csl
header-includes:
  \usepackage{caption}
  \usepackage{setspace}\doublespacing
  # \usepackage{lineno} \linenumbers
---

@@

pdf(paste0(figdir,"methodsfig.pdf"), bg="transparent", height = 13, width = 12)

@@

plotlims <- c(0.5,4.5,0,15)
ticklocsx <- 1:4    # locations of tick marks on x axis
ticklocsy <- seq(0,15,5)    # locations of tick marks on y axis
ticklabelsx <- c('sc','mc','lc','mel')        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

coneab <- coneab[c(1:3,5),]   # don't plot rod predictions
for (b in 1:4){
  polygon(c(b-0.3,b,b,b-0.3),c(0,0,coneab[b,2],coneab[b,2]),col=colpal[2],border=NA)
  polygon(c(b+0.3,b,b,b+0.3),c(0,0,coneab[b,3],coneab[b,3]),col='black',border=NA)
}

text(4.5,14,'(d)',pos=4,cex=2)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Wavelength (nm)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Difference power", col.lab=rgb(0,0,0), line=2.8, cex.lab=1.5)
title(main='Luminance',cex.main=2)

@@

pcontrast <- pcontrast[c(1,2,3,5),]   # don't plot rods

plotlims <- c(0.5,4.5,-100,100)
ticklocsx <- 1:4    # locations of tick marks on x axis
ticklocsy <- seq(-100,100,50)    # locations of tick marks on y axis
ticklabelsx <- c('sc','mc','lc','mel')        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:4){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,4],pcontrast[b,4]),col=stimcolours[1],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,5],pcontrast[b,5]),col=stimcolours[2],border='black',lwd=2)
}

text(3.65,80,'(i)',pos=4,cex=2)

plotlims <- c(0.5,4.5,-20,20)
ticklocsx <- 1:4    # locations of tick marks on x axis
ticklocsy <- seq(-20,20,10)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:4){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,6],pcontrast[b,6]),col=stimcolours[3],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,7],pcontrast[b,7]),col=stimcolours[4],border='black',lwd=2)
}

text(3.6,18,'(j)',pos=4,cex=2)

plotlims <- c(0.5,4.5,-50,50)
ticklocsx <- 1:4    # locations of tick marks on x axis
ticklocsy <- seq(-50,50,25)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:4){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,8],pcontrast[b,8]),col=stimcolours[5],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,9],pcontrast[b,9]),col=stimcolours[6],border='black',lwd=2)
}

text(3.6,45,'(k)',pos=4,cex=2)

plotlims <- c(0.5,4.5,-20,20)
ticklocsx <- 1:4    # locations of tick marks on x axis
ticklocsy <- seq(-20,20,10)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:4){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,10],pcontrast[b,10]),col=stimcolours[7],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,11],pcontrast[b,11]),col=stimcolours[8],border='black',lwd=2)
}

@@

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.5Hz (mm)", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main='Luminance',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.125,c('Monocular','Binocular','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.14,'(a)',adj=0.5,cex=2.5)

@@

legend(-24,0.6,c('Prior','Lum 1F', 'L-M 1F', 'S 1F', 'Mel 1F'), col=c('black',stimcolours[c(1,4,6,7)]),lwd=c(2,6,6,6,6),lty=1, cex=1.8, box.lwd=2, bg='white')

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.4Hz (mm)", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main='Luminance',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmasksP[condlist[cond],]),contrastsdB,cleanmasksPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP[condlist[cond],]),contrastsdB,cleanmasksPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksP[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.125,c('Monocular','Binocular cross','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.14,'(f)',adj=0.5,cex=2.5)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims2[1:2], ylim=plotlims2[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy2, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy2, side = 2, at=ticklocsy2, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 1Hz (mm)", col.lab=rgb(0,0,0), line=3.5, cex.lab=1.5)
title(main='Luminance',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum2[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP2[condlist[cond],]),contrastsdB,cleanmeansP2CI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP2[condlist[cond],]),contrastsdB,cleanmeansP2CI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP2[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.0125,c('Monocular','Binocular','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.014,'(a)',adj=0.5,cex=2.5)

@@

legend(-24,0.6,c('Prior','Lum 2F', 'L-M 2F', 'S 2F', 'Mel 2F'), col=c('black',stimcolours[c(1,4,6,7)]),lwd=c(2,6,6,6,6),lty=1, cex=1.8, box.lwd=2, bg='white')

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims2[1:2], ylim=plotlims2[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy2, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy2, side = 2, at=ticklocsy2, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.8Hz (mm)", col.lab=rgb(0,0,0), line=3.5, cex.lab=1.5)
title(main='Luminance',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum2[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmasksP2[condlist[cond],]),contrastsdB,cleanmasksP2CI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP2[condlist[cond],]),contrastsdB,cleanmasksP2CI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksP2[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.0125,c('Monocular','Binocular cross','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.014,'(f)',adj=0.5,cex=2.5)

@@

legend(-0.019,0.0155,c('Luminance','L-M','S','Melanopsin'),lty=1,col=stimcolours[c(1,3,6,7)],lwd=3,pch=21,pt.cex=2,pt.bg=stimcolours[c(2,4,5,8)],cex=1.1,box.lwd=3,bg='white')

@@

Pupil diameters are regulated by the autonomic nervous system, which combines light signals across the eyes independently of the visual cortex. Distinct classes of retinal photoreceptor are involved in this process, with cones and rods driving the initial constriction and intrinsically photosensitive retinal ganglion cells maintaining diameter over prolonged time periods. We investigated binocular combination by targeting different photoreceptor pathways using a novel binocular multiprimary system to modulate the input spectra via silent substitution. At the first harmonic of the modulation frequency, luminance and S-cone responses showed strong binocular facilitation, and weak interocular suppression. Melanopsin responses were invariant to the number of eyes stimulated. Notably, the L-M pathway involved binocular inhibition, whereby responses to binocular stimulation were weaker than for monocular stimulation. The second harmonic involved strong interocular suppression in all pathways, but with some evidence of binocular facilitation. Our results are consistent with a computational model of binocular signal combination (implemented in a Bayesian hierarchical framework), in which the weight of interocular suppression differs across pathways. We also find pathway differences in response phase, consistent with different lag times for phototransduction. This work demonstrates for the first time the algorithm governing binocular combination in the autonomic nervous system.

@@

Pupil diameters are determined by a network of subcortical nuclei that must combine incoming light levels across the two eyes, independently of cortical visual areas. We investigated the algorithmic properties of this binocular combination process for the first time. Using silent subsitution, we targeted different post-retinal pathways (luminance, L-M cones, S-cones, and melanopsin) with flickering stimuli generated by a novel binocular multi-primary system. We found different patterns of binocular combination across pathways, and also differences in the phase of the pupil response. The results were well described by a computational model of binocular signal combination, in which the weight of suppression between the eyes differed across pathways. This elucidates the algorithm governing binocular combination in the autonomic nervous system.

@@

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex [@Hubel1962]. For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye [@Baker2018; @Campbell1965]. At higher contrasts, a process of 'ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal [@Baker2007; @Moradi2009]. This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work [@Segala2023] showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli directed at the luminance, L-M cone, S-cone, and melanopsin pathways (see Figure \@ref(fig:spectraplots)e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision [@Meese2006] implemented within a hierarchical Bayesian framework.

@@

knitr::include_graphics(paste0(figdir,'methodsfig.pdf'))

@@

>>>>

---
title: 'Binocular combination in the autonomic nervous system'
author: |
 | Federico G. Segala$^1$, Aurelio Bruno$^{1,2}$, Joel T. Martin$^{1,3}$,
 | Anisa Y. Morsi$^1$, Alex R. Wade$^{1,4}$ & Daniel H. Baker$^{1,4}$
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: true
  html_document: default
bibliography: references.bib
link-citations: yes
linkcolor: black
csl: the-journal-of-neuroscience.csl
header-includes:
  \usepackage{caption}
  \usepackage{setspace}\doublespacing
  \usepackage{lineno} \linenumbers
  \newcommand{\beginsupplement}{\setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}}
---

@@

pdf(paste0(figdir,"methodsfigrods.pdf"), bg="transparent", height = 13, width = 12)

@@

plotlims <- c(0.5,5.5,0,15)
ticklocsx <- 1:5    # locations of tick marks on x axis
ticklocsy <- seq(0,15,5)    # locations of tick marks on y axis
ticklabelsx <- c('sc','mc','lc','rods','mel')        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

# coneab <- coneab[c(1:3,5),]   # don't plot rod predictions
for (b in 1:5){
  polygon(c(b-0.3,b,b,b-0.3),c(0,0,coneab[b,2],coneab[b,2]),col=colpal[2],border=NA)
  polygon(c(b+0.3,b,b,b+0.3),c(0,0,coneab[b,3],coneab[b,3]),col='black',border=NA)
}

text(5.5,14,'(d)',pos=4,cex=2)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Wavelength (nm)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Difference power", col.lab=rgb(0,0,0), line=2.8, cex.lab=1.5)
title(main='Light flux',cex.main=2)

@@

# pcontrast <- pcontrast[c(1,2,3,5),]   # don't plot rods

plotlims <- c(0.5,5.5,-100,100)
ticklocsx <- 1:5    # locations of tick marks on x axis
ticklocsy <- seq(-100,100,50)    # locations of tick marks on y axis
ticklabelsx <- c('sc','mc','lc','rods','mel')        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:5){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,4],pcontrast[b,4]),col=stimcolours[1],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,5],pcontrast[b,5]),col=stimcolours[2],border='black',lwd=2)
}

lines(c(0.5,5.5),c(0,0),lwd=2)

text(4.65,80,'(i)',pos=4,cex=2)

plotlims <- c(0.5,5.5,-20,20)
ticklocsx <- 1:5    # locations of tick marks on x axis
ticklocsy <- seq(-20,20,10)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:5){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,6],pcontrast[b,6]),col=stimcolours[3],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,7],pcontrast[b,7]),col=stimcolours[4],border='black',lwd=2)
}

lines(c(0.5,5.5),c(0,0),lwd=2)

text(4.6,18,'(j)',pos=4,cex=2)

plotlims <- c(0.5,5.5,-50,50)
ticklocsx <- 1:5    # locations of tick marks on x axis
ticklocsy <- seq(-50,50,25)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:5){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,8],pcontrast[b,8]),col=stimcolours[5],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,9],pcontrast[b,9]),col=stimcolours[6],border='black',lwd=2)
}

lines(c(0.5,5.5),c(0,0),lwd=2)

text(4.6,45,'(k)',pos=4,cex=2)

plotlims <- c(0.5,5.5,-20,20)
ticklocsx <- 1:5    # locations of tick marks on x axis
ticklocsy <- seq(-20,20,10)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

@@

for (b in 1:5){
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,10],pcontrast[b,10]),col=stimcolours[7],border='black',lwd=2)
  polygon(c(b-0.3,b+0.3,b+0.3,b-0.3),c(0,0,pcontrast[b,11],pcontrast[b,11]),col=stimcolours[8],border='black',lwd=2)
}

lines(c(0.5,5.5),c(0,0),lwd=2)

@@

layout(matrix(c(1,2,3,4,5,5,5,5,6,7,8,9),6,2,byrow=TRUE))

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.5Hz (mm)", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main='Light flux',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.125,c('Monocular','Binocular','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.14,'(a)',adj=0.5,cex=2.5)

@@

legend(-24,0.6,c('Prior','LF 1F', 'L-M 1F', 'S 1F', 'Mel 1F'), col=c('black',stimcolours[c(1,4,6,7)]),lwd=c(2,6,6,6,6),lty=1, cex=1.8, box.lwd=2, bg='white')

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.4Hz (mm)", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main='Light flux',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmasksP[condlist[cond],]),contrastsdB,cleanmasksPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP[condlist[cond],]),contrastsdB,cleanmasksPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksP[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.125,c('Monocular','Binocular cross','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.14,'(f)',adj=0.5,cex=2.5)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims2[1:2], ylim=plotlims2[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy2, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy2, side = 2, at=ticklocsy2, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 1Hz (mm)", col.lab=rgb(0,0,0), line=3.5, cex.lab=1.5)
title(main='Light flux',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum2[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP2[condlist[cond],]),contrastsdB,cleanmeansP2CI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP2[condlist[cond],]),contrastsdB,cleanmeansP2CI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP2[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.0125,c('Monocular','Binocular','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.014,'(a)',adj=0.5,cex=2.5)

@@

legend(-24,0.6,c('Prior','LF 2F', 'L-M 2F', 'S 2F', 'Mel 2F'), col=c('black',stimcolours[c(1,4,6,7)]),lwd=c(2,6,6,6,6),lty=1, cex=1.8, box.lwd=2, bg='white')

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims2[1:2], ylim=plotlims2[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy2, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     # add the tick labels
mtext(text = ticklabelsy2, side = 2, at=ticklocsy2, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 0.8Hz (mm)", col.lab=rgb(0,0,0), line=3.5, cex.lab=1.5)
title(main='Light flux',cex.main=2)
for (cond in 1:3){
  lines(contrastsfinedB,modrespsLum2[condlist2[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)
  # lines(contrastsfinedB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmasksP2[condlist[cond],]),contrastsdB,cleanmasksP2CI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP2[condlist[cond],]),contrastsdB,cleanmasksP2CI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksP2[condlist[cond],]), pch = 20+cond, col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(15,0.0125,c('Monocular','Binocular cross','Dichoptic cross'),pch=21:23,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2,cex=1.2)
text(17,0.014,'(f)',adj=0.5,cex=2.5)

@@

legend(-0.019,0.0155,c('Light flux','L-M','S','Melanopsin'),lty=1,col=stimcolours[c(1,3,6,7)],lwd=3,pch=21,pt.cex=2,pt.bg=stimcolours[c(2,4,5,8)],cex=1.1,box.lwd=3,bg='white')

@@

Pupil diameters are regulated by the autonomic nervous system, which combines light signals across the eyes independently of the visual cortex. Distinct classes of retinal photoreceptor are involved in this process, with cones and rods driving the initial constriction and intrinsically photosensitive retinal ganglion cells maintaining diameter over prolonged time periods. We investigated binocular combination by targeting different photoreceptor pathways using a novel binocular multiprimary system to modulate the input spectra via silent substitution. At the first harmonic of the modulation frequency, light flux and S-cone responses showed strong binocular facilitation, and weak interocular suppression. Melanopsin responses were invariant to the number of eyes stimulated. Notably, the L-M pathway involved binocular inhibition, whereby responses to binocular stimulation were weaker than for monocular stimulation. The second harmonic involved strong interocular suppression in all pathways, but with some evidence of binocular facilitation. Our results are consistent with a computational model of binocular signal combination (implemented in a Bayesian hierarchical framework), in which the weight of interocular suppression differs across pathways. We also find pathway differences in response phase, consistent with different lag times for phototransduction. This work demonstrates for the first time the algorithm governing binocular combination in the autonomic nervous system.

@@

Pupil diameters are determined by a network of subcortical nuclei that must combine incoming light levels across the two eyes, independently of cortical visual areas. We investigated the algorithmic properties of this binocular combination process for the first time. Using silent subsitution, we targeted different post-retinal pathways (light flux, L-M cones, S-cones, and melanopsin) with flickering stimuli generated by a novel binocular multi-primary system. We found different patterns of binocular combination across pathways, and also differences in the phase of the pupil response. The results were well described by a computational model of binocular signal combination, in which the weight of suppression between the eyes differed across pathways. This elucidates the algorithm governing binocular combination in the autonomic nervous system.

@@

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex [@Hubel1962]. For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye [@Baker2018; @Campbell1965]. At higher contrasts, a process of 'ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal [@Baker2007; @Moradi2009]. This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work [@Segala2023] showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli directed at the light flux, L-M cone, S-cone, and melanopsin pathways (see Figure \@ref(fig:spectraplots)e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision [@Meese2006] implemented within a hierarchical Bayesian framework.

@@

knitr::include_graphics(paste0(figdir,'methodsfigrods.pdf'))

@@

# Supplementary material

\beginsupplement

```{r primarylinearity, fig.cap="Linearity of primaries for each device. The calibration measures were summed (i.e., total unweighted irradiance), and the input-output relationship summarised by a 7th order polynomial reverse curve fit. By applying the coefficients of the regression, it is possible to achieve a linear output in the primaries.", fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'SuppFig1.pdf'))
